---
title: "flexdashboard: Easy interactive dashboards for R"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source: embed
---

# Intro text {.sidebar}

You can use flexdashboard to publish groups of related data visualizations as a dashboard. A flexdashboard can either be static (a standard web page) or dynamic (a Shiny interactive document). A wide variety of components can be included in flexdashboard layouts, including:

```{r include = FALSE}

knitr::opts_chunk$set(cache = TRUE)

library(dplyr)
library(highcharter)
library(viridisLite)
library(ggplot2)
data(unemployment)
data(uscountygeojson)


thm <- 
  hc_theme(
    colors = c("#1a6ecc", "#434348", "#90ed7d"),
    chart = list(
      backgroundColor = "transparent",
      style = list(fontFamily = "Source Sans Pro")
    ),
    xAxis = list(
      gridLineWidth = 1
    )
  )

data("USArrests", package = "datasets")
data("usgeojson")

USArrests <- USArrests %>%
  mutate(state = rownames(.))

n <- 4
colstops <- data.frame(
  q = 0:n/n,
  c = substring(viridis(n + 1), 0, 7)) %>%
  list.parse2()
```

# Page 1 {data-icon="fa-list"}

## Column {data-height=350}

### ggplot2 Plots

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
```

## Column {data-height=650 .tabset .tabset-fade}

### htmlwidgets {data-width=300}

```{r}
highchart() %>%
  hc_add_series_map(usgeojson, USArrests, name = "Sales",
                    value = "Murder", joinBy = c("woename", "state"),
                    dataLabels = list(enabled = TRUE,
                                      format = '{point.properties.postalcode}')) %>%
  hc_colorAxis(stops = colstops) %>%
  hc_legend(valueDecimals = 0, valueSuffix = "%") %>%
  hc_mapNavigation(enabled = TRUE) %>%
  hc_add_theme(thm)
```

### tables {data-width=700}

```{r}
library(DT)
datatable(iris, options = list(pageLength = 10))
```

# Page 2 {data-icon="fa-map"}

## Column

### base plots

```{r}
plot(mpg$displ, mpg$hwy)
```

> A more exhaustive list includes: Interactive JavaScript data visualizations based on htmlwidgets; R graphical output including base, lattice, and grid graphics; tabular data (with optional sorting, filtering, and paging); value boxes for highlighting important summary data; gauges for displaying values on a meter within a specified range; and text annotations of various kinds.
