---
output: 
  html_document
params:
  symbol: AMZN
---
  
# Report for `r params$symbol` on `r Sys.Date()`

## Summary

```{r setup, echo = FALSE, message = FALSE}

library(quantmod)
library(DT)
library(dygraphs)
library(forecast)
library(highcharter)

prices <- round(getSymbols(params$symbol, src = 'google', auto.assign = FALSE), 2)
close <- Cl(last(prices))
open <- Op(last(prices))

recent <- last(prices, n=90)
recent_nv <- recent[,-5]

```

The stock closed `r ifelse(close>open,'up','down')` at `r close` dollars per share yesterday.

### Price History

The chart below is made with the `quantmod` and `highcharter` R packages. An API returns all of the price history based on the stock tick symbol provided as a parameter. The candlestick chart is a default function from highcharter, as is the the Economist theme.


```{r echo=FALSE}

highchart() %>% 
   hc_yAxis_multiples(
    list(title = list(text = NULL), height = "75%", top = "0%"),
    list(title = list(text = NULL), height = "15%", top = "80.5%", opposite = TRUE)
  ) %>%
  hc_add_series(prices, yAxis = 0, name = params$symbol) %>% 
  hc_add_series(prices[,paste0(params$symbol,".Volume")], name = "Volume", type = "column", yAxis = 1) %>% 
  hc_add_theme(hc_theme_economist())
```

## Model

```{r, warning = FALSE, message = FALSE}
m <- arima(recent[,1], c(3,1,2))
f <- forecast(m,7)
```

The forecast for tomorrow is `r round(as.numeric(f$mean)[1],2)` dollars.

This model is fit with the arima function in the forecast package. A 3,1,2 ARIMA model is used though in practice any range of models could be used.

```{r, warning = FALSE, message = FALSE}
plot(forecast(m,7), main = "")
```


