---
output: 
  html_document
---

## Report for GOOG as of `r Sys.Date()`

```{r setup, echo=FALSE, message=FALSE}

library(quantmod)
library(DT)
library(dygraphs)
library(forecast)
library(highcharter)


prices <- round(getSymbols("GOOG", src = 'google', auto.assign = FALSE), 2)
close <- Cl(last(prices))
open <- Op(last(prices))

recent <- last(prices, n = 90)
recent_nv <- recent[, -5]

```

The stock closed `r ifelse(close>open,'up','down')` at `r close` dollars per share yesterday.

### Price History

The chart below is made with the `quantmod` and `highcharter` R packages. An API returns all of the price history based on the stock tick symbol provided as a parameter. The candlestick chart is a default function from highcharter, as is the the Economist theme.


```{r echo = FALSE}

highchart() %>% 
   hc_yAxis_multiples(
    list(title = list(text = NULL), height = "75%", top = "0%"),
    list(title = list(text = NULL), height = "15%", top = "80.5%", opposite = TRUE)
  ) %>%
  hc_add_series(prices, yAxis = 0, name = "GOOG") %>% 
  hc_add_series(prices[,paste0("GOOG",".Volume")], name = "Volume", type = "column", yAxis = 1) %>% 
  hc_add_theme(hc_theme_economist())
```


## Model

```{r, warning = FALSE, message = FALSE}
m <- arima(recent[,1], c(3,1,2))
f <- forecast(m,7)
```

The forecast for tomorrow is `r round(as.numeric(f$mean)[1],2)` dollars.

This model is fit with the arima function in the forecast package. A 3,1,2 ARIMA model is used though in practice any range of models could be used.

```{r, warning = FALSE, message = FALSE}
plot(forecast(m,7), main = "")
```


