---
output: html_document
---

# `r params$symbol`

```{r echo=FALSE, message=FALSE}
# set up

library(xts)
library(quantmod)

prices <- getSymbols(params$symbol, auto.assign = FALSE)
move <- Cl(last(prices)) - Op(last(prices))
```

## Recommendation: `r ifelse(move > 0, "BUY", "SELL")`

`r params$symbol` will `r ifelse(move > 0, "increase", "decrease")` in price during the next trading period. 


## Price History

The chart below is made with the `quantmod` R package, a widely used package for collecting and visualizing financial data in R. You can learn more about `quantmod` at the website [www.quantmod.com](rstudio.github.io/DT/).  

***

```{r echo=FALSE}
chartSeries(prices, theme = chartTheme("white", bg.col = "white"))
```

## Method

This forecast was predicted with the *recency algorithm*, a simple---probably useless---method for determining stock prices. The recency algorithm predicts that the next price movement, $M_{j}$, will be in the same direction as the most recent price movement. $M_{i}$.

$$M_{i}= Close_{i} - Open_{i}$$

$$
    M_{j}= 
\begin{cases}
    > 0,              & \text{if } M_{i} > 0\\
    \leq 0,              & \text{if } M_{i}\leq 0
\end{cases}
$$

## Raw Data

The table below displays the daily price data for the stock. In the next section, we will learn how to make a concise, interactive table with the `DT` package, a new package for making searchable data tables. You can learn more about the `DT` package at the website [rstudio.github.io/DT/](rstudio.github.io/DT/).  

***

```{r echo=FALSE, comment=''}
as.data.frame(prices)[1:5 , 1:4]
```

