---
title: "Packrat"
subtitle: "A Dependency Management System for R"
author: "Kevin Ushey (kevin@rstudio.com)"
date: "February 3rd, 2016"
output:
  ioslides_presentation: default
css: styles.css
---

## Packrat

Packrat is an R package that helps you manage your project's R package dependencies in a way that is:

> * **Reproducible** <p style="margin-top: 10px;">Packrat records the exact package versions you depend on, and ensures those exact versions are the ones that get installed wherever you go.</p>

> * **Isolated** <p style="margin-top: 10px;">Installing a new or updated package for one project won't break your other projects, and vice versa. That's because packrat gives each project its own private package library.</p>

> * **Portable** <p style="margin-top: 10px;">Easily transport your projects from one computer to another, even across different platforms. Packrat makes it easy to install the packages your project depends on.</p>

## Reproducibility

Packrat helps you to save and preserve the environment in which an R project or analysis is run in.

Ever finished an analysis, come back a year later, and asked:

> * _Who wrote this garbage code?_ <span style="display: inline-block; font-size: 9pt; margin-bottom: 12px;">(oops, it was me)</span>

> * _Why is `lattice` throwing an error when I make figures now? Apparently its API changed?_

> * _Okay, I re-ran the analysis, but now `nlme` is complaining about model convergence: I swear this didn't happen before..._

> * _What happened to that R package we got from our collaborators? I swear it's in my e-mail somewhere..._

## Reproducibility {.build}

Packrat helps you solve this problem by:

> - **Ensuring package sources are stored alongside your project,**

> - **Tracking package versions so that, if you need to roll back a package for some reason, you can do so**

Packrat lets you build a sandbox in which an analysis can run, and helps you ensure that sandbox can persist over time.

## Isolation {.build}

With the current state of the world, when you install an R package, it gets installed into a **global** library -- all projects on your system have access to that same library ...

... but only one version of a particular package can be installed at a time.

This is problematic if you have older projects that, for example, depend on older versions of `Rcpp`, but want to leverage features of newer `Rcpp` for other projects.

We need some mechanism of **isolating** projects, and their dependencies, from each other -- and Packrat helps us do this.

## Isolation {.build}

Packrat makes your projects more isolated by giving each project its own **private library**.

A `library` is just a folder of R packages -- when you make a call to the `library` function, e.g.

    library(dplyr)
    
R attempts to load `dplyr` from your user, or system, library.

If you use Packrat to manage an R project, it will tell R to use and load packages from the project's **private library** instead.

This also means that calls to `install.packages`, `devtools::install_github`, `biocLite` and others will install into the **private library** by default -- no need to worry about clobbering packages in your user library!

## Portability {.build}

Because Packrat collects and tracks all project dependencies, it ensures that the environment used for a particular analysis is **portable** across different systems.

If you are working with collaborators, you can ensure that everyone working on a Packrat project is using the **same version** of any R packages your project depends on.

# Using Packrat {.segue}

## Main Packrat Verbs {.build}

Packrat maintains two sets of state:

> - The **snapshot** of your project's R dependencies, and

> - The **private library** powering your project.

Packrat has two main verbs for interacting with these two components:

> `packrat::snapshot()` records the package versions used by a project and downloads their source code for storage with the project,

> `packrat::restore()` applies the previous snapshot to a directory (building packages from source as necessary).

Use `packrat::snapshot()` to **save** state, and use `packrat::restore()` to **load** state.

## Demo

These are concepts much more easily explored with an interactive demo, so let's start!

## Packrat Fundamentals

```{r, eval=FALSE}
packrat::init()
```

Create a packrat project within a directory, giving the project its own private package library.

```{r, eval=FALSE}
packrat::snapshot()
```

Finds the packages in use in the project and stores a list of those packages, their current versions, and their *source code*.

```{r, eval=FALSE}
packrat::restore()
```

Restore the directory to the last snapshotted state (building packages from source as necessary).


## Initializing a Project

```{r, eval=FALSE}
packrat::init()
```

```
Adding these packages to packrat:
            _          
    packrat   0.4.0.12

Fetching sources for packrat (0.4.0.12) ... OK (GitHub)
Snapshot written to '~/projects/reshape/packrat/packrat.lock'
Installing packrat (0.4.0.12) ... OK (built source)
Initialization complete!
```

## Snapshotting Installed Packages

```{r, eval=FALSE}
packrat::snapshot()
```

```
Adding these packages to packrat:
             _       
    plyr       1.8.1 
    Rcpp       0.11.2
    reshape2   1.4   
    stringr    0.6.2 

Fetching sources for plyr (1.8.1) ... OK (CRAN current)
Fetching sources for Rcpp (0.11.2) ... OK (CRAN current)
Fetching sources for reshape2 (1.4) ... OK (CRAN current)
Fetching sources for stringr (0.6.2) ... OK (CRAN current)
Snapshot written to '~/projects/reshape/packrat/packrat.lock'
```

## Restoring the State of the Library

```{r, eval=FALSE}
packrat::restore()
```

```
Installing Rcpp (0.11.2) ... OK (downloaded binary)
Installing stringr (0.6.2) ... OK (downloaded binary)
Installing plyr (1.8.1) ... OK (downloaded binary)
Installing reshape2 (1.4) ... OK (downloaded binary)
```

## Updating a Package from Github

```{r, eval=FALSE}
packrat::install_github("RcppCore/Rcpp")
```

```{r, eval=FALSE}
packrat::snapshot()
```

```
Upgrading these packages already present in packrat:
             from         to
    Rcpp   0.11.2   0.11.2.1

Snapshot written to '~/projects/reshape/packrat/packrat.lock'
```

```{r, eval=FALSE}
packrat::restore()
```

```
Installing Rcpp (0.11.2.1) ... OK (built source)
```

## Bundling and Unbundling

```{r, eval=FALSE}
packrat::bundle()
```

```
The packrat project has been bundled at:
- "~/projects/reshape/packrat/bundles/reshape-2014-06-24.tar.gz"
```

```{r, eval=FALSE}
packrat::unbundle("reshape-2014-06-24.tar.gz", where = "~/Desktop")
```

```
- Untarring 'reshape-2014-06-24.tar.gz' in directory '~/Desktop'...
- Restoring project library...
Installing packrat (0.4.0.12) ... OK (built source)
Installing Rcpp (0.11.2.1) ... OK (built source)
Installing stringr (0.6.2) ... OK (downloaded binary)
Installing plyr (1.8.1) ... OK (downloaded binary)
Installing reshape2 (1.4) ... OK (downloaded binary)
Done! The project has been unbundled and restored at:
- "~/Desktop/reshape"
```

## Packrat Objectives

* Isolated, portable, and reproducible environment for R projects

* Capture all source code required to reproduce configurations

* Flexible and easy to use solution to the problem of reproducibility:
    - "One button" snapshot/restore
    - Simple and convenient archiving (bundle/unbundle)
    - Optional integration with version control

* Foundation for even higher fidelity reproducibility
    - e.g. Creating VMs or Docker containers for projects


## Questions?

Packrat website: <http://rstudio.github.io/packrat>

Packrat source: <https://github.com/rstudio/packrat>
