---
title: "An R Markdown Document"
author:
- Li Lei
- Han Meimei
date: "2015/10/28"
output:
  ioslides_presentation: default
  beamer_presentation:
    theme: AnnArbor
  slidy_presentation: default
bibliography: references.bib
---

## Start with a cool section

You can use traditional **Markdown** syntax, such as [links](http://yihui.name/knitr) and `code`. Here is a quote:

> A girl phoned me the other day and said "Come on over, there's nobody home." I went over. Nobody was home. -- Rodney Dangerfield

## Followed by another section

Of course you can write lists:

- apple
- pear
- banana

Or ordered lists:

1. items
1. will
1. be
1. ordered
    - nested
    - items

## Okay, some R code

```{r linear-model}
fit = lm(dist ~ speed, data = cars)
b = coef(fit)  # coefficients
summary(fit)
```

The code will be highlighted in all output formats.

## And some pictures

```{r lm-vis, fig.cap='Regression diagnostics'}
par(mfrow = c(2, 2), pch = 20, mar = c(4, 4, 2, .1), bg = 'white')
plot(fit)
```

## A little bit math

Our regression equation is $Y=`r b[1]`+`r b[2]`x$, and the model is:

$$ Y = \beta_0 + \beta_1 x + \epsilon$$

## Pandoc extension: definition lists

Programmer
:   A programmer is the one who turns coffee into code.
LaTeX
:   A simple tool that is nothing but a couple of backslashes.

## Pandoc extension: examples

We have some examples.

(@) Think what is `0.3 + 0.4 - 0.7`. Zero. Easy.
(@weird) Now think what is `0.3 - 0.7 + 0.4`. Still zero?

People are often surprised by (@weird).

## Pandoc extension: tables

A table here.

Table:  Demonstration of simple table syntax.

  Right     Left     Center     Default
-------     ------ ----------   -------
     12     12        12            12
    123     123       123          123
      1     1          1             1

## Pandoc extension: footnotes

We can also write footnotes[^1].

[^1]: hi, I'm a footnote

Or write some inline footnotes^[as you can see here].

## Pandoc extension: citations

We compile the R Markdown file to Markdown through **knitr** [@xie2013] in R [@R-base]. For more about @xie2013, see <http://yihui.name/knitr>.

## References
